Forecasting the old-age dependency ratio to determine the best pension age

### 2020-09-11 PensionAge.Rmd
Finished section 2

In this paper, we use the OADR as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working-age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07]. The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure. Forecasts of population structure have been built on forecasts of age- and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age- and gender-specific population structure.

@@

The Lee-Carter (LC) method has been widely used to forecast mortality rates across various countries, including Australia. The model proposed by @LC92 is
\begin{equation}\label{eq:lc}
  \log(m_{x,t})=a_x+b_xk_t+\varepsilon_{x,t},
\end{equation}
where $m_{x,t}$ is the central death rate (properly defined in Section \@ref(sec:data)) for age $x$ in year $t$. This model involves fitting a matrix of log central death rates with sets of age-specific constants, {$a_x$} and {$b_x$}, and a time-varying index $k_t$. $b_x$ is a set of parameters that tell us the rate at which each age responds to changes in $k_t$. $k_t$ measures the general level of the log death rates; $a_x$ is the general pattern across the age of the log mortality rate and $\varepsilon_{x,t}$ is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is over-parameterised. For example, if the vector $a,b,k$ is one solution, then for any scalar $c$, $a, b/c, kc$ or $a - bc, b, k+c$ also must be two possible solutions. Therefore,  the constraints
\begin{equation}
  \sum_{t=1}^{n}k_t=0, \qquad\text{and}\qquad \sum_{x=x_1}^{x_m}b_x=1
\end{equation}
are imposed, and a singular value decomposition is used to find the least squares solution of \eqref{eq:lc} [@Good1969]. The estimates of $b_x$ and $k_t$ are therefore the first principal component and scores of $\bm{X}$ with $(i,t)$th element $[\log(m_{x_i,t})-a_{x_i}]$, where $a_{x_i}$ is the sample mean of $\log(m_{x_i,t})$ over $t$. The log transformation introduces some bias, so the LC method adjusts $k_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $k_t$ is then extrapolated using a random walk with drift model, which is equivalent to an ARIMA ($0,1,0$) model with a drift parameter.

@@

When we apply the @HU07 model to data on males and females, we need to constrain the resulting forecasts so they do not diverge [@HBY13]. In the product-ratio method, we apply the HU model to the geometric mean of male and female log mortality data, and the ratio of male and female log mortality data, rather than to each series independently. That is, we define
\begin{align}
  p_t(x)&=\sqrt{s_{t}^{\text{M}}(x)s_{t}^{\text{F}}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t}^{\text{M}}(x)}{s_{t}^{\text{F}}(x)}}. \label{eq:coherent_2}
\end{align}
Then, the HU method is applied to these two new variables instead of male and female series:
\begin{align*}
  \log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  \log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with an ARIMA model while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA model (i.e., $d$ in ARIMA($p,d,q$) constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecasts of male and female do not diverge. Then, forecasts of sex-specific data are obtained using
\begin{align*}
  m_{n+h\mid n}^{\text{M}}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n}^{\text{F}}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for forecasting net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration, and a sum and difference is used rather than a product and ratio.

We use the same notation and similar procedures to construct data as in @HB08. The data used to construct population age structure included: age- and gender-specific birth and death numbers of each calendar year, age- and gender-specific population members on 1\ January of each year, and age- and gender-specific exposures to risk (i.e., the population of age $x$ at 30 June) for each year. The notations are as follows:
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$}\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$}\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$}\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$}
\end{align*}
where $x=0,1,2,\dots,p-1,p^+$ denotes age with $p^+$ the open-ended upper age group, set to 100$^+$ in this study. The years is denoted by $t=1,2,\dots,n$, and superscripts M and F denote male and female respectively.

@@

@@

## Forecasting population age structure and OADR {#sec:forecasting}

In this study, three steps were taken to simulate OADR associated with a specific pension age scheme. A stochastic functional data model was applied to five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates to simulate these five components over the prediction horizon (which is 50 years in this study). Age- and gender-specific population paths over the prediction horizon were simulated using the demographic growth-balance equation. The Monte Carlo simulation method was used consistently in this research. Finally, calculation of OADR associated with a specific pension age scheme was performed by applying the OADR formula. Even though the process seems complicated, it is relatively easy to implement using \textit{R} [@Team18]. The detailed explanation of each step is as follows. The corresponding R code is attached as a supplement.

### Step 1 {-}

The main procedure for forecasting the population age structure follows @HB08, except that the data are updated, and we incorporate the product-ratio method [@HBY13] to improve the accuracy of the forecast.

, introduced in section \@ref(sec:HU), is applied to the mortality rate, fertility rate and net migration separately, based on the assumption that they behave independently. Therefore, to fit the model to historical fertility rates, let $y_t(x)$ in equation \eqref{eq:HU_1} denote the observed historical fertility data.

### Step 2 {-}

After simulating paths for age-specific fertility rate, age- and gender-specific mortality rate and age- and gender-specific net migration, future population value was simulated using the demographic growth-balance equations over the prediction horizon. The future deaths and births in this equation are assumed to follow a Poisson distribution, with parameters as a function of future mortality and fertility rates (see Hyndman and Booth, 2008 for the details of the functions). Hence, we can simulate $h$-step-ahead age- and gender-specific births and deaths based on simulated mortality and fertility rates. According to the demographic growth-balance equation, one-step-ahead net migrations, deaths, births and population in the last year were simulated, allowing for the generation of one-step-ahead population predictions by sex and age. This simulation was repeated for years $t=n+1,\dots,n+h$ to obtain a path of population. The process was repeated and simulated $i$ paths of population based on Monte Carlo simulation method were created (we set $i=1000$ in this research). It should be noted that there are three types of randomness in the simulation: randomness from the second equation in the Hyndman-Ullah specification (i.e., $e_t(x)$ in \eqref{eq:HU_1}), randomness from the ARIMA model fitted to principal component scores and randomness from the Poisson distribution. It should be noted that the simulated population will be an array with three dimensions: age, year and simulation.

### Step 3 {-}

@@

Figures \@ref(fig:totalpopulationbyyears) and \@ref(fig:totalpopulationin2034) show the forecast results of population. Figure \@ref(fig:totalpopulationbyyears) shows the forecast of the total population by gender over the prediction horizon, while Figure \@ref(fig:totalpopulationin2028) shows the forecast of population age structure in a specific year, 2038. Figure \@ref(fig:totalpopulationbyyears) indicates that both the female and male population are expected to increase to around 27 million by 2068. Since the fertility rate is expected to remain low, the possible reason for the significant increase in population size is net migration. We can also observe that the further the projected year is, the bigger the forecast variation of a population. This increase in variation shows that we have less relevant information for further apart prediction horizons.

>>>>

In this paper, we use the OADR as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working-age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07]. The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure. Forecasts of population structure have been built on forecasts of age- and sex-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age- and sex-specific population structure.

@@

The Lee-Carter (LC) method has been widely used to forecast mortality rates across various countries, including Australia. The model proposed by @LC92 (but using our notation) is
\begin{equation}\label{eq:lc}
  \log m_{t}(x)=\mu(x)+\phi(x)\beta_{t}+\varepsilon_{t,x},
\end{equation}
where $m_{t}(x)$ is the central death rate (properly defined in Section \@ref(sec:data)) for age $x$ in year $t$. This model involves fitting a matrix of log central death rates with sets of age-specific constants, {$\mu(x)$} and {$\phi(x)$}, and a time-varying index $\beta_{t}$. $\phi(x)$ is a set of parameters that tell us the rate at which each age responds to changes in $\beta_{t}$. $\beta_{t}$ measures the general level of the log death rates; $\mu(x)$ is the general pattern across the age of the log mortality rate and $\varepsilon_{t,x}$ is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is over-parameterised. Therefore, the constraints
\begin{equation}
  \sum_{t=1}^{n}\beta_t=0, \qquad\text{and}\qquad \int_x \phi(x)=1
\end{equation}
are imposed, and a singular value decomposition is used to find the least squares solution of \eqref{eq:lc} [@Good1969]. The estimates of $\phi(x)$ and $\beta_t$ are therefore the first principal component and scores of $\bm{X}$ with $(i,t)$th element $[\log(m_{t}(x))-\mu(x)]$, where $\mu(x)$ is the sample mean of $\log(m_{t}(x))$ over $t$. The log transformation introduces some bias, so the LC method adjusts $\beta_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $\beta_t$ is then extrapolated using a random walk with drift model, which is equivalent to an ARIMA ($0,1,0$) model with a drift parameter.

@@

When we apply the @HU07 model to data on males and females, we need to constrain the resulting forecasts so they do not diverge. In the product-ratio method [@HBY13], we apply the HU model to the geometric mean of male and female log mortality data, and the ratio of male and female log mortality data, rather than to each series independently. That is, we define
\begin{align}
  p_t(x)&=\sqrt{s_{t}^{\text{M}}(x)s_{t}^{\text{F}}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t}^{\text{M}}(x)}{s_{t}^{\text{F}}(x)}}. \label{eq:coherent_2}
\end{align}
Then, the HU method is applied to these two new variables instead of male and female series:
\begin{align*}
  \log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  \log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with an ARIMA model while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA model (i.e., the order of differencing is constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecasts of male and female do not diverge. Then, forecasts of sex-specific data are obtained using
\begin{align*}
  m_{n+h\mid n}^{\text{M}}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n}^{\text{F}}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for forecasting net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration, and a sum and difference is used rather than a product and ratio.

## Stochastic population forecasting

We follow the general procedure proposed in @HB08, except that we use the product-ratio method rather than the HU method for mortality and net-migration.

The data used are age- and sex-specific birth and death numbers of each calendar year, age- and sex-specific population members on 1&nbsp;January of each year, and age- and sex-specific exposures to risk (i.e., the population of age $x$ at 30&nbsp;June) for each year. Let
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$},\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$},\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$},\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$},
\end{align*}
where $x=0,1,2,\dots,p-1,p^+$ denotes age with $p^+$ the open-ended upper age group, set to 100$^+$ in this study. The year is denoted by $t=1,2,\dots,n$, and superscripts $M$ and $F$ denote male and female respectively.

@@

A stochastic functional data model was applied to the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates to simulate these five components into the future.

We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration. From these models, sample paths are simulated for age-specific fertility rates, age- and sex-specific mortality rates and age- and sex-specific net migration. Then future population values are simulated using the demographic growth-balance equations. The future deaths and births in this equation are assumed to follow a Poisson distribution, with parameters as a function of future mortality and fertility rates. We bootstrap the residuals of the net-migration model to simulate future net-migration numbers (see @HB08 for the details). Hence, we can simulate $h$-step-ahead age- and sex-specific population numbers based on simulated births, deaths and net-migrants.

This simulation was repeated for years $t=n+1,\dots,n+h$ to obtain a sample path of the age-specific population. Many simulated population paths were obtained in this way.

It should be noted that there are three types of randomness in the simulation: randomness from the second equation in the Hyndman-Ullah specification (i.e., $e_t(x)$ in \eqref{eq:HU_1}), randomness from the ARIMA model fitted to principal component scores, and randomness from the Poisson distribution for births and deaths or the bootstrap for migrants.

@@

## Forecasting OADR {#sec:forecasting}

@@

Figures \@ref(fig:totalpopulationbyyears) and \@ref(fig:totalpopulationin2034) show the forecast results of population. Figure \@ref(fig:totalpopulationbyyears) shows the forecast of the total population by sex over the prediction horizon, while Figure \@ref(fig:totalpopulationin2028) shows the forecast of population age structure in a specific year, 2038. Figure \@ref(fig:totalpopulationbyyears) indicates that both the female and male population are expected to increase to around 27 million by 2068. Since the fertility rate is expected to remain low, the possible reason for the significant increase in population size is net migration. We can also observe that the further the projected year is, the bigger the forecast variation of a population. This increase in variation shows that we have less relevant information for further apart prediction horizons.
